---
title: "Lab5"
author: "Casey Zhang"
date: "2024-10-03"
output: "github_document"
---
# Load packages and dataset
```{r}
library(haven)
library(ggplot2)
library(Rmisc)
library(rstatix)
library(emmeans)
library(bruceR)
library(dplyr)
library(labelled)

lab5 <- read.csv("/Users/caseyzhang/Documents/GitHub/Lab5/lab5data.csv")

```

# Recode Variables
```{r}
lab5 <- lab5 %>%
  mutate_at(c('Gender'),funs(str_replace(., "0", "Men")))

lab5 <- lab5 %>%
  mutate_at(c('Gender'),funs(str_replace(., "1", "Women")))

lab5 <- lab5 %>%
  mutate_at(c('Greek'),funs(str_replace(., "0", "No")))

lab5 <- lab5 %>%
  mutate_at(c('Greek'),funs(str_replace(., "1", "Yes")))


```
# Note that for this lab you will assume assumptions are met, but for your own dataset, you should always check assumptions first (Not for this lab, but for the data anlaysis project you should check your assumptions)

# How to conduct planned contrasts?
## Step 1: Recode 2x2 (greek*gender) into 1x4 (non-greek women, greek women, non-greek men, greek men)
```{r}

lab5$Group <- ifelse(lab5$Gender == "Women" & lab5$Greek == "Yes", "Greek Women", 
                     ifelse(lab5$Gender == "Men" & lab5$Greek == "Yes", "Greek Men", 
                            ifelse(lab5$Gender == "Women" & lab5$Greek == "No", "Non-Greek Women", 
                                   "Non-Greek Men")))


```

## Step 2: Write down contrast coefficients
```{r}

#Here we have 4 groups. 

#Because of the order you coded above, it would assume that 
#1st group is Greek Women, 
#2nd group is Greek men, 
#3rd group is Non-Greek Women,
#4th group is Non-Greek Men. 

#So if you want to compare Greek Women vs. Greek Men you would write (-1, 1, 0, 0)
# For groups you aren't comparing, put 0; arbitrarily assign greek women=-1, greek men=1

```

## Step 3: Build the model and test it
```{r}

model <- lm(Drinks ~ Group, data = lab5)

emm<- emmeans(model, "Group")

emm

contrast(emm, list("Greek Women vs. Greek Men" = c(-1, 1, 0, 0)))

contrast(emm, list("Greek Main Effect" = c(1, 1, -1, -1)))

contrast(emm, list("Interaction" = c(1, -1, -1, 1)))


#Note that even though you can still test main effects and interaction using the contrast method, but if all you're interested is main effects and interaction, the method below is probably easier.

#Own note: result is that Greek men drink more than greek women (mean is larger) and it's statistically significant (p>0.05)

```

# How to test 2x2, main effects and interaction then simple effects
```{r}

#Please note here that the t statistics may look slightly different than the contrast method above, but the overall interpretation shouldn't change

#Also please note that this method is okay, but you're sort of thinking about what the results mean after seeing the full results, whereas for contrasts you need to understand what you're testing first before running the test

mod<-MANOVA(lab5, dv = "Drinks", between = c("Gender", "Greek")) 

EMMEANS(mod, effect = "Gender", by = "Greek", p.adjust = "none")

EMMEANS(mod, effect = "Greek", by = "Gender", p.adjust = "none")

#Since we already coded 2x2 into 1x4, below is another way to test the simple effects, but it's basically post-hocs. I did not use p value adjustment here, but you can change it to bonferroni or something else for your own data.
#MANOVA(b, dv = "Drinks", between = c("Group")) %>%
  #EMMEANS("Group", p.adjust = "none")

#Own notes
    #Top table 1: descriptive statistics, top table 2: main effects and interactions
    #Gender: Men drink more than women (p<.05, mens average is higher than women)
    #Greek: Greeks drinks more
    #Levene's test of homogeneity of variance (it's significant so it's not equal) -> correct the data by transforming it (log)
    #Simple effect of gender: for non-greek people, drinking isn't that different between genders
    #Pairwise comparisons of "greek": compairing (yes-no=greek vs. non greek) for men (greek vs. non-grekn) vs. women (greek vs.     non-greek)
```
# Visualize your results using bar graphs
```{r}
#Bar graphs are different than the graphs you made before. You first need to computed the mean scores and standard errors for bar graphs.

lab5_clean <- lab5 %>%
  drop_na(Drinks)

plot<-summarySE(lab5_clean, measurevar="Drinks", groupvars=c("Gender", "Greek"))

plot

plot2<-summarySE(lab5_clean, measurevar="Drinks", groupvars=c("Group"))

plot2

ggplot(plot, aes(x = Greek, y = Drinks, fill = Greek)) +
  geom_col() + facet_wrap(~ Gender) + theme_bruce()

ggplot(plot2, aes(x = Group, y = Drinks, fill = Group)) +
  geom_col()  + theme_bruce() + theme(axis.text.x = element_text(angle = -10))

#3rd graph: 2*2 graph
#4th graph: 1*4 graph

```

# Q1: You hypothesized that Greek Women have a lower GPA than non-Greek women. Conduct the correct analysis below and explain whether or not your hypothesis is supported. 

Collect GPA from the women and conduct a simple effect comparison between the GPA of greek women and non-greek women. The hypothesis is not supported as the p-value is greater than the standard threshold of 0.05. This means that greek women and non-greek women do not have statistically significantly different GPAs.
```{r}

model <- lm(GPA ~ Group, data = lab5)

emm<- emmeans(model, "Group")

emm

contrast(emm, list("Greek Women vs. Non-Greek Women" = c(-1, 0, 1, 0)))

```

# Q2: You hypothesized that on average women have a higher GPA than men. Conduct the correct analysis below and explain whether or not your hypothesis is supported. 

If you only look at the main effect it doesn't give you the full picture so we need to look at both the main effect of gender and the simple effects of gender in the context of Greek life and gender in the context of non-Greek life. The results show that there is no statistically significant main effect nor simple effects of any of the conditions as the p-values are all above the threshold of 0.05. This means that on average, women do not have a higher GPA than men. This applies when comparing genders within Greek life and outside of Greek life.

```{r}

model <- lm(GPA ~ Group, data = lab5)

emm<- emmeans(model, "Group")

emm

#Main effect:
contrast(emm, list("Gender Main Effect" = c(-1, 1, -1, 1)))

#Simple effect:
contrast(emm, list("Greek Women vs. Greek Men" = c(-1, 1, 0, 0)))
contrast(emm, list("Non-Greek Women vs. Non-Greek Men" = c(0, 0, -1, 1)))


```

# Q3: Create a bar graph to compare GPA by gender and greek (either graph works)
```{r}

lab5_clean <- lab5 %>%
  drop_na(GPA)

#Notw: drop_na is dropping the missing values (taking out that row)

plot<-summarySE(lab5_clean, measurevar="GPA", groupvars=c("Gender", "Greek"))

plot

plot2<-summarySE(lab5_clean, measurevar="GPA", groupvars=c("Group"))

plot2

ggplot(plot, aes(x = Greek, y = GPA, fill = Greek)) +
  geom_col() + facet_wrap(~ Gender) + theme_bruce()

ggplot(plot2, aes(x = Group, y = GPA, fill = Group)) +
  geom_col()  + theme_bruce() + theme(axis.text.x = element_text(angle = -10))

```
